---
title: "Suicide rate"
date: 2019-05-14T10-24-00
authors: ["admin"]
output: 
  blogdown::html_page:
    toc: true
summary: Analysis of suicide rate worldwide since 1985.
categories: ["visualization", "R"]
tags: ["ggplot2", "gganimate", "gif"]
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.width=12, fig.height=8,
  cache = T
)

```
# Intro

A data about suicide rate was analyzed in this project. Mostly I will focus on visualization by various graphs. As someone famous said: "There are no good analysis, only useful ones", with that in mind, my illustrations are up to reader's interpretation. Furthur analysis into each country will be made in the near future.

Data were taken from:

United Nations Development Program. (2018). Human development index (HDI). Retrieved from http://hdr.undp.org/en/indicators/137506

World Bank. (2018). World development indicators: GDP (current US$) by country:1985 to 2016. Retrieved from http://databank.worldbank.org/data/source/world-development-indicators#

[Szamil]. (2017). Suicide in the Twenty-First Century [dataset]. Retrieved from https://www.kaggle.com/szamil/suicide-in-the-twenty-first-century/notebook

World Health Organization. (2018). Suicide prevention. Retrieved from http://www.who.int/mental_health/suicide-prevention/en/

# Load library ------------------

```{r, warning = F, message = F}
library(tidyverse)
library(skimr)
library(maps)
library(gganimate)
library(ggrepel)
library(maps)

theme_set(theme_minimal() +
            theme(panel.background = element_blank(),
                  plot.title = element_text(size = 28,
                                            face = "bold",
                                            color = "#222222"),
                  plot.subtitle = ggplot2::element_text(size=22,
                                                        margin=ggplot2::margin(9,0,9,0)),
                  legend.text.align = 0,
                  legend.background = ggplot2::element_blank(),
                  legend.key = ggplot2::element_blank(),
                  legend.text = ggplot2::element_text(size=10,
                                                      color="#222222"),
                  axis.text = ggplot2::element_text(size=10,
                                                    color="#222222"),
                  axis.text.x = ggplot2::element_text(margin=ggplot2::margin(5, b = 10),
                                                      vjust = 0.5, size = 10),
                  axis.text.y = ggplot2::element_text(margin=ggplot2::margin(5, b = 10), 
                                                      vjust = 0.5, size = 10)))
```

# Load data ------------------

```{r message=FALSE}
data <- read_csv("data/master.csv")
data <- data %>% rename(HDI = `HDI for year`,
                        suicides_rate = `suicides/100k pop`,
                        gdp_yearly = `gdp_for_year ($)`,
                        gdp_capita = `gdp_per_capita ($)`)
data$age <- factor(data$age, levels = c("5-14 years", "15-24 years", 
                                        "25-34 years", "35-54 years", 
                                        "55-74 years", "75+ years"))
```

# Load extra data for worldmaps and statistics

## Load world map

```{r}
world <- map_data("world")
```

## Draw world map

```{r}

world[world$region == "Antigua" | world$region == "Barbuda",]$region <- "Antigua and Barbuda"
world[world$region == "Cape Verde",]$region <- "Cabo Verde"
world[world$region == "South Korea",]$region <- "Republic of Korea"
world[world$region == "Russia",]$region <- "Russian Federation"
world[world$region == "Saint Kitts" | world$region == "Nevis",]$region <- "Saint Kitts and Nevis"
world[world$region == "Saint Vincent" | world$region == "Grenadines",]$region <- "Saint Vincent and Grenadines"
world[world$region == "Trinidad" | world$region == "Tobago",]$region <- "Trinidad and Tobago"
world[world$region == "UK",]$region <- "United Kingdom"
world[world$region == "USA",]$region <- "United States"
world[world$subregion == "Macao" & !is.na(world$subregion),]$region <- "Macau"

worldmap <- ggplot(data = world, aes(long, lat, group = group)) + geom_polygon(fill = "#f2f2f2") +
  theme(panel.background = element_blank(),
                   axis.title = element_blank(),
                   axis.line.x = element_blank(),
                   axis.ticks = element_blank(),
                   axis.text = element_blank()) +
   coord_fixed(1.2)

```



## Add EU

```{r message=FALSE}

eu <- read_csv("data/listofeucountries.csv") %>% pull(x)

eu <- replace(eu, eu == "Slovak Republic", "Slovakia")

```


# Data Exploratory ----------------------- 

## Have a look

```{r}
summary(data)
```

- country-year column is just a concatenation of country and year column

- suicides/100k pop is calculated by suicides / population * 100000

- HDI is Human Development Report, seems missing alot, the only variable is missing

- Age is divided into brackets

## A closer look:

```{r}

skim_with(numeric = list(hist = NULL))
skim(data) 

```

There are:

- 6 Age brackets

- 101 differenet countries in this data set

- Year from 1985 to 2016. (32 years)

- 2321 combinations of country-year (less than 32 * 101). Must be some implicit missing data with year and country

- 6 different generations

- 2 Sex

## Extract country stats

HDI, GDP per year and per capital are values based on country so it would make sense to extract those values into another 
dataframe

```{r}
countrystat <- data %>% select(country, year, gdp_yearly, gdp_capita, HDI) %>% 
  distinct()

data <- data %>% select(-gdp_yearly, -gdp_capita, -HDI, -`country-year`)

```


## Review implicit missing data

```{r}
data <- data %>% complete(country, year, sex, age)
```

Pattern of missing data

```{r}
data %>% group_by(country, year) %>% 
  summarise(miss = sum(is.na(suicides_no))) %>% 
  ungroup() %>% count(miss)
```


So each country every year can either miss all data, have all data, but there are some country only miss 2 data, Let's review those

```{r}
data %>% filter(year == 2016) %>% right_join(data %>% group_by(country, year) %>% 
  summarise(miss = sum(is.na(suicides_no))) %>% 
  filter(!miss %in% c(0,12)), by = c("country", "year")) %>% 
  filter(is.na(suicides_no))
```

Data from 2016 are missing with age group 5-14 years old

Let's find out how many missing data with each country each year

```{r}
data %>% group_by(country, year) %>% summarise(avg_rate = mean(suicides_rate)) %>% summarise(n = sum(is.na(avg_rate))) %>% arrange(desc(n))
  
```

Plot on map


```{r fig.height=8, fig.width=12}

worldmap + data %>% 
  group_by(country, year) %>% 
  summarise(n = sum(is.na(suicides_rate))/12) %>%
  ungroup()  %>% 
  group_by(country) %>% 
  summarise(n = sum(n)) %>% 
  left_join(world, by = c("country" = "region")) %>%  
  geom_polygon(data = ., aes(fill = n))  + 
  scale_fill_viridis_c(trans = "log2", name = "Years of missing data", 
                       guide = guide_colorbar(title.position = "top", 
                                              direction = "horizontal",
                                              nbin = 10)) +
  labs(title = "Missing data map", subtitle = "Since 1985") +
  theme(legend.position = "bottom",
        legend.justification = "center",
        legend.key.width = unit(3,"cm"),
        panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(), 
        legend.title.align = 0.5,
        legend.title = element_text(size = 20, face = "bold", color = "#222222"))
```

Most of the data are from Russia, America nand Europe. There are some countries that have a lot of missing data

Missing pattern
```{r, fig.height= 12}
data %>% 
  group_by(country, year) %>% 
  summarise(na = sum(is.na(suicides_no))) %>% 
  filter(na > 0) %>% 
  ggplot(aes(year, country, group = country)) + 
  geom_tile() + 
  theme( panel.grid = element_blank()) +
  scale_x_continuous(expand = c(0,0)) + 
  labs(title = "Pattern of missing data")
```

There are a lot of random missing data through out the years

## What is generation's age range?

```{r, message = F, warning = F}
data %>% filter(!is.na(generation)) %>% 
  separate(age, into =c("min", "max")) %>% mutate(max = ifelse(max== "years", 100, max)) %>% 
  mutate(min = as.integer(min), max = as.integer(max)) %>% 
  mutate(min = year - min, max = year - max) %>% 
  ggplot() + geom_jitter(aes(generation, min, color = "Max possible birth year"), alpha = 0.3) +
  geom_jitter(aes(generation, max, color = "Min possible birth year"), alpha = 0.3) +
  scale_color_manual(name = "", values = c("green", "red")) + theme(legend.position = "top") + 
  labs(title = "Birth year of all generations", y = "Birth year", x = "Generations")
```

- Since Age values are not provided but put in the age range, we can only estimate the actual year birth.

From the data, birth year of :

- G.I generation is aroung 1900

- Silent generation is around 1925

- Boomers generation is around 1955

- Generation X is around 1975

- Millennials is around 1980

- Generation Z is around 2000

Compare to Wiki

- G.I generation birth year is from 1900s to late 1920s

- Silent generation birth year is from late 1920s to mid 1940s

- Boomers generation birth year is from 1946 to 1964

- Genration X birth year is from early-to-mid 1960s to the early 1980s

- Millennials birth year is from early 1980s to early 2000s

- Generation Z birth year is from 1990s till now

There is no big discrepancy between data set and Wiki, no outliner either, so it is safe to assume that there is no mistake in our data.

# Analyze

## Suicides rate by year all over the world

Set up data

```{r}
rate <- data %>% 
  group_by(country, year) %>% 
  summarise(suicides_rate = sum(suicides_no)/sum(population) * 1e5) %>% 
  ungroup()
```


### Suicides rate per country per year of all recorded country visualized in boxplot.
The number on top shows number of countries recorded in each year.


```{r}
rate %>%
  filter(!is.na(suicides_rate)) %>% group_by(year) %>% mutate(n = n()) %>% 
  ggplot(aes(x = factor(year))) +
  geom_boxplot(aes(y = suicides_rate )) + 
  geom_text(aes(label = n, y = 55)) +
  theme(axis.text.x = element_text(angle = 45)) +
  labs(title = "Distribution of suicide rate of all country", subtitle = "Since 1985" , y = "Suicide rate", x = "Year")

```

There is no noticable trend

### Rate of suicides per country

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

rate %>% 
  ggplot(aes(year, suicides_rate, color = country)) + 
  geom_line() +
  geom_text_repel(data = . %>% ungroup() %>%
                    group_by(year) %>% 
                    top_n(n = 5, wt = suicides_rate) %>%
                    ungroup() %>% 
                    complete(country, year),
                  aes(label = country), hjust = 0,  
                  segment.alpha = 0.2, xlim = c(2015,NA)) +
  coord_cartesian(clip = "off") +
  theme(legend.position = "none",
        plot.margin = margin(1,4,1,1, "cm")) +
  labs(title = "Rate of suicides", subtitle = "in {round(frame_along)}", x = "Year", y = "Suicide rate") +
  transition_reveal(year) 


```

### Top recorded suicide rate in each year with barplot

```{r message=FALSE, warning=FALSE}
(rate %>% ungroup() %>%
  group_by(year) %>% 
  top_n(n = 5, wt = suicides_rate) %>% 
  mutate(rank = rank(-suicides_rate)) %>% 
  ggplot(aes(rank, suicides_rate, fill = country)) + 
  geom_col() +
  geom_label(aes(label = country ),fill = "white") +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 20),
        axis.title.x = element_text(size = 20)) +
  scale_x_reverse() +
  coord_flip() +
  transition_states(year, transition_length = 2, state_length = 2) +
  labs(subtitle = "in {closest_state}", title = "Top recorded suicide rate", y = "Suicide rate")+
  enter_drift(x_mod = -1) + exit_drift(x_mod = 1)) %>% 
  animate(nframes = 300, height = 800, width = 800 )

```

### Recorded suicides rate in each year with map

```{r fig.height=8, fig.width=12, message=FALSE, warning=FALSE}
(worldmap +
   geom_polygon(data = rate %>%
                  left_join(world, by = c("country" = "region")) %>% 
                  filter(!is.na(suicides_rate)),
                aes(fill = suicides_rate))  + 
   scale_fill_viridis_c(name = "Suicides rate", 
                       guide = guide_colorbar(title.position = "top", 
                                              direction = "horizontal"),
                       na.value  = "#f2f2f2") +
  theme(legend.position = "bottom",
        legend.key.width = unit(5,"cm"),
        legend.title = element_text(size = 20, face = "bold", color = "#222222"),
        legend.title.align = 0.5,
        panel.grid = element_blank(),
        plot.margin = margin(15,1,1,1),
        axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
  labs(subtitle = "in {closest_state}", title = "Suicide rate worldwide")+
  transition_states(year)) %>% 
  animate(duration = 20, height = 800, width = 1200)

```




### Top 10 biggest increase in suicide rate in 1 year all time

```{r}
rate %>% 
  mutate(lag = suicides_rate - lag(suicides_rate)) %>% 
  top_n(10, lag) %>% 
  arrange(desc(lag))
```

### Top 10 biggest decrease in suicide rate in 1 year

```{r paged.print=FALSE}
rate %>% 
  mutate(lag = suicides_rate - lag(suicides_rate)) %>% 
  top_n(10, -lag) %>% 
  arrange(lag)
```

Based on this, more information can be obtained to get a further insight of the events in these countries

### Biggest changes

```{r message=FALSE, warning=FALSE}

(ggplot(data = rate %>% filter(country %in% (rate %>% group_by(country) %>% 
  filter(!is.na(suicides_rate)) %>% 
  summarise(delta = max(suicides_rate, na.rm = T) - min(suicides_rate, na.rm = T)) %>% 
  arrange(desc(delta)) %>% 
  top_n(10, delta) %>% pull(country))),aes(year, suicides_rate, color = country) ) + 
  geom_line() +
  geom_text_repel(aes(label = country), xlim = c(2020, NA), hjust = 0, segment.alpha = 0.2) +
  coord_cartesian(clip = "off") +
  theme(legend.position = "none",
        plot.margin = margin(1, 3.5, 1, 1, "cm")) +
  labs(title = "Top changes in rate of suicide", subtitle = "in {round(frame_along)}", x = "Year", y = "Suicide rate") +
  transition_reveal(year)) %>% 
  animate(duration = 20)

```

### Suicide rate by GDP

```{r message=FALSE, warning=FALSE}
rate %>% left_join(countrystat) %>% 
  ggplot(aes(suicides_rate, gdp_capita)) + 
  geom_point() +
  labs(y = "GDP per capita", x = "Suicide rate", title = "Suicide rate by GDP", subtitle = "Since 1985")

```

There is no indication of a relationship between them

## Number of suicides per country

Prepare data

```{r}
n_suicides <- data %>% 
  group_by(country,year) %>% 
  summarise(n = sum(suicides_no))
```

### Number of suicides per country each year

```{r, warning=FALSE, message=FALSE}

n_suicides %>% 
  ggplot(aes(year, n, color = country)) +
  geom_line() +  
  geom_text_repel(data = . %>% ungroup() %>%
                    group_by(year) %>% 
                    top_n(n = 5, wt = n) %>%
                    ungroup() %>% 
                    complete(country, year), 
                  aes(label = country ), xlim = c(2015,NA), hjust = 0,
                  segment.alpha = 0.2) + 
  coord_cartesian(clip = 'off') +
  theme(legend.position = "none",  
        plot.margin = margin(1,4,1,1, "cm")) +
  labs(title = "Number of suicides", subtitle = "in {round(frame_along)}", x = "Year", y = "Number of rate") +
  transition_reveal(year)

```

## From 2010

Total number of suicide since 2010

```{r fig.height=8, fig.width=12}
worldmap + data %>% filter(year >= 2010 & year < 2016) %>%
  group_by(country) %>% 
  summarise(n = sum(suicides_no, na.rm = T)) %>%
  filter(!is.na(n)) %>%
  left_join(world, by = c("country" = "region")) %>%  
  geom_polygon(data = ., aes(fill = n))  + 
  scale_fill_viridis_c(name = "Number of suicides",
                       trans = "log10",
                       guide = guide_colorbar(title.position = "top", 
                                              direction = "horizontal")) +
  labs(title = "Total number of suicide", subtitle = "Since 2010") +
  theme(legend.position = "bottom",
        legend.key.width = unit(5,"cm"),
        legend.title = element_text(size = 20, face = "bold", color = "#222222"),
        legend.title.align = 0.5,
        panel.grid = element_blank(),
        plot.margin = margin(15,1,1,1),
        axis.text.x = element_blank(),
        axis.text.y = element_blank())
```

### Average rate of suicide since 2010

```{r fig.height=8, fig.width=12}
worldmap + rate %>% filter(year >= 2010 & year < 2016) %>%
  group_by(country) %>% summarise(n = mean(suicides_rate, na.rm = T)) %>% filter(!is.na(n)) %>% 
  left_join(world, by = c("country" = "region")) %>%  
  geom_polygon(data = ., aes(fill = n))  + 
  scale_fill_viridis_c(name = "Average number of suicides",
                       guide = guide_colorbar(title.position = "top", 
                                              direction = "horizontal")) +
  labs(title = "Average rate of suicide", subtitle = "Since 2010") +
  theme(legend.position = "bottom",
        legend.key.width = unit(5,"cm"),
        legend.title = element_text(size = 20, face = "bold", color = "#222222"),
        legend.title.align = 0.5,
        panel.grid = element_blank(),
        plot.margin = margin(15,1,1,1),
        axis.text.x = element_blank(),
        axis.text.y = element_blank())
```

## Gender

### Number of suicides by gender each year in the world

```{r}
data %>% group_by(year, sex) %>% 
  summarise(suicides_no = sum(suicides_no, na.rm = T)) %>% 
  ggplot(aes(year, suicides_no, color = sex)) + 
  geom_line() + 
  theme(legend.position = "none") +
  labs(title = "Number of suicide worldwide", x = "Year", y = "Suicide number", subtitle = "since 1985")
```

### Rate of suicides by gender each year in the world

```{r}
data %>% group_by(year, sex) %>% 
  summarise(suicides_no = sum(suicides_no, na.rm = T)/ sum(population, na.rm = T) * 1e5) %>% 
  ggplot(aes(year, suicides_no, color = sex)) + 
  geom_line() + 
  theme(legend.position = "none") +
  labs(title = "Rate of suicide worldwide", x = "Year", y = "Suicide rate", subtitle = "since 1985")
```

### Number of suicides by gender each year in every country

```{r}
data %>% group_by(country, year, sex) %>% 
  summarise(suicides_no = sum(suicides_no)) %>% 
  ggplot(aes(year, suicides_no, color = country)) + 
  geom_line() + facet_wrap(~sex) +
  theme(legend.position = "none") +
  labs(title = "Number of suicides by gender", subtitle = "Since 1985", x = "Year",
       y = "Number of suicide")

```


### Rate of suicides by gender each year in every country

```{r}
data %>% group_by(country, year, sex) %>% 
  summarise(suicides_rate = sum(suicides_no)/sum(population) * 1e5) %>% 
  ggplot(aes(year, suicides_rate, color = country)) + 
  geom_line() + facet_wrap(~sex) +
  theme(legend.position = "none") +  
  labs(title = "Rate of suicides by gender", subtitle = "Since 1985", x = "Year",
       y = "Rate of suicide")
  

```

## Age group

### Number of suicides each age group in the world

```{r}
data %>% group_by(year, age) %>% 
  summarise(suicides_no = sum(suicides_no, na.rm = T)) %>% 
  ggplot(aes(year, suicides_no, color = fct_reorder2(age, year, suicides_no))) + 
  geom_line() +
  scale_color_discrete(name = "Age group") +
  labs(title = "Number of suicides each age group", subtitle = "Since 1985", x = "Year",
       y = "Number of suicides")
```

Number of suicides in 35-54 age group is the biggest due to large population proportion, in the next graph, age group 75+ actually has the highest suicide rate

### Rate of suicides each age group in the world

```{r}
data %>% group_by(year, age) %>% 
  summarise(suicides_rate = sum(suicides_no, na.rm = T)/sum(population, na.rm = T) * 1e5) %>% 
  ggplot(aes(year, suicides_rate, color = fct_reorder2(age, year, suicides_rate))) + 
  geom_line() +
  scale_color_discrete(name = "Age group") +
  labs(title = "Rate of suicides each age group", subtitle = "Since 1985", x = "Year",
       y = "Rate of suicides")

```


### Number of suicides by age group each year in every country

```{r}
data %>% group_by(country, year, age) %>% 
  summarise(suicides_no = sum(suicides_no)) %>% 
  ggplot(aes(year, suicides_no, color = country)) + 
  geom_line() +
  geom_text_repel(data = data %>% 
              group_by(country, year, age) %>% 
              summarise(suicides_no = sum(suicides_no)) %>% 
              filter(year == 2000) %>% 
              ungroup() %>% 
              group_by(age, year) %>% 
              top_n(3,suicides_no),
            aes(label = country))+
  facet_wrap(~ age) +
  theme(legend.position = "none") +
  labs(title = "Number of suicides by age group", subtitle = "Since 1985",
       x = "Year", y = "Number of suicides")

```

### Rate of suicides by age group each year in every country

```{r}
data %>% group_by(country, year, age) %>% 
  summarise(suicides_rate = sum(suicides_no)/sum(population)*1e5) %>% 
  ggplot(aes(year, suicides_rate, color = country)) + 
  geom_line() + facet_wrap(~ age) +
  theme(legend.position = "none") +
  labs(title = "Rate of suicides by age group", subtitle = "Since 1985",
       x = "Year", y = "Rate of suicides")

```

### Rate of suicides by age group and sex

```{r fig.height=12, fig.width=8}
data %>% group_by(country, year, age, sex) %>% 
  summarise(suicides_rate = sum(suicides_no)/sum(population)*1e5) %>% 
  ggplot(aes(year, suicides_rate, color = country)) + 
  geom_line() + facet_grid(age ~ sex)+
  theme(legend.position = "none") +
  labs(title = "Number of suicides by age and sex", subtitle = "Since 1985",
       x = "Year", y = "Number of suicides")
```


## EU

EU background map

```{r}
EUmap <- ggplot(data = world %>% 
                  filter(region %in% eu), aes(long, lat, group = group)) + geom_polygon(fill = "#f2f2f2") +
  theme(panel.background = element_blank(),
                   axis.title = element_blank(),
                   axis.line.x = element_blank(),
                   axis.ticks = element_blank(),
                   axis.text = element_blank()) +
   coord_fixed(1.2)

```

Extract EU data

```{r}
EUrate <- data %>% 
  filter(country %in% eu) %>% 
  group_by(country, year) %>% 
  summarise(suicides_rate = sum(suicides_no)/sum(population) * 1e5) %>% 
  ungroup()
```

### Suicides rate per country per year of all recorded country visualized in boxplot.
The number on top shows number of countries recorded in each year.


```{r}
EUrate %>%
  filter(!is.na(suicides_rate)) %>% group_by(year) %>% mutate(n = n()) %>% 
  ggplot(aes(x = factor(year))) +
  geom_boxplot(aes(y = suicides_rate )) + 
  geom_text(aes(label = n, y = 0.0006)) +
  geom_line(data = rate %>% 
              group_by(year) %>% 
              summarise(suicides_rate = median(suicides_rate, na.rm = T)), 
            aes(x= factor(year), y = suicides_rate, group = "The World"), color = "red", size = 1) +
  geom_text(aes(factor(2010), 7, label = "The world median"), color = "red", size = 4) +
  theme(axis.text.x = element_text(angle = 45)) +
  labs(title = "Distribution of suicide rate of all country", subtitle = "Since 1985 in EU" , y = "Suicide rate", x = "Year")

```

EU has higher suicides rate than the world


### Rate of suicides per country

```{r message=FALSE, warning=FALSE}

EUrate %>% 
  ggplot(aes(year, suicides_rate, color = country)) + 
  geom_line() +
  geom_text_repel(data = . %>% ungroup() %>%
                    group_by(year) %>% 
                    top_n(n = 5, wt = suicides_rate) %>%
                    ungroup() %>% 
                    complete(country, year),
                  aes(label = country), hjust = 0,  
                  segment.alpha = 0.2, xlim = c(2015,NA)) +
  coord_cartesian(clip = "off") +
  theme(legend.position = "none",
        plot.margin = margin(1,4,1,1, "cm")) +
  labs(title = "Rate of suicides in EU", subtitle = "in {round(frame_along)}", x = "Year", y = "Suicide rate") +
  transition_reveal(year) 


```

### Top recorded suicides rate in each year with barplot

```{r message=FALSE, warning=FALSE}
(EUrate %>% ungroup() %>%
  group_by(year) %>% 
  top_n(n = 5, wt = suicides_rate) %>% 
  mutate(rank = rank(-suicides_rate)) %>% 
  ggplot(aes(rank, suicides_rate, fill = country)) + 
  geom_col() +
  geom_label(aes(label = country ),fill = "white") +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 20),
        axis.title.x = element_text(size = 20)) +
  scale_x_reverse() +
  coord_flip() +
  transition_states(year, transition_length = 2, state_length = 2) +
  labs(subtitle = "in {closest_state}", title = "Top recorded suicide rate", y = "Suicide rate") +
  enter_drift(x_mod = 6) + exit_drift(x_mod = -1)) %>% 
  animate(nframes = 300, height = 800, width = 800 )

```

### Recorded suicides rate in each year with map

```{r fig.height=8, fig.width=12, message=FALSE, warning=FALSE}
(EUmap +
   geom_polygon(data = EUrate %>%
                  left_join(world, by = c("country" = "region")) %>% 
                  filter(!is.na(suicides_rate)),
                aes(fill = suicides_rate))  + 
   scale_fill_viridis_c(name = "Suicides rate", 
                       guide = guide_colorbar(title.position = "top", 
                                              direction = "horizontal"),
                       na.value  = "#f2f2f2") +
  theme(legend.position = "bottom",
        legend.key.width = unit(5,"cm"),
        legend.title = element_text(size = 20, face = "bold", color = "#222222"),
        legend.title.align = 0.5,
        panel.grid = element_blank(),
        plot.margin = margin(15,1,1,1),
        axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
  labs(subtitle = "in {closest_state}", title = "Suicide rate in EU") +
  transition_states(year)) %>% 
  animate(duration = 20, height = 800, width = 1200)

```




### Top 10 biggest increase in suicide rate in 1 year all time

```{r}
EUrate %>% 
  mutate(lag = suicides_rate - lag(suicides_rate)) %>% 
  top_n(10, lag) %>% 
  arrange(desc(lag))
```

### Top 10 biggest decrease in suicide rate in 1 year

```{r paged.print=FALSE}
EUrate %>% 
  mutate(lag = suicides_rate - lag(suicides_rate)) %>% 
  top_n(10, -lag) %>% 
  arrange(lag)
```

Based on this, more information can be obtained to get a further insight of the events in these countries

### Biggest changes

```{r message=FALSE, warning=FALSE}

(ggplot(data = EUrate %>% 
          filter(country %in% 
                   (EUrate %>% group_by(country) %>%
                      filter(!is.na(suicides_rate)) %>% 
                      summarise(delta = max(suicides_rate, na.rm = T) - min(suicides_rate, na.rm = T)) %>% 
                      arrange(desc(delta)) %>% 
                      top_n(5, delta) %>% pull(country))),
        aes(year, suicides_rate, color = country)) + 
  geom_line() +
  geom_text_repel(aes(label = country), xlim = c(2020, NA), hjust = 0, segment.alpha = 0.2) +
  coord_cartesian(clip = "off") +
  theme(legend.position = "none",
        plot.margin = margin(1, 3.5, 1, 1, "cm")) +
  labs(title = "Top changes in rate of suicide", subtitle = "in {round(frame_along)} in EU", x = "Year", 
       y = "Suicide rate") +
  transition_reveal(year)) %>% 
  animate(duration = 20)


```

### Suicide rate by GDP

```{r message=FALSE, warning=FALSE}
EUrate %>% left_join(countrystat) %>% 
  ggplot(aes(suicides_rate, gdp_capita)) + 
  geom_point() +
  labs(y = "GDP per capita", x = "Suicide rate", title = "Suicide rate by GDP in EU", subtitle = "Since 1985")

```

There is no indication of a relationship between them

### Rate of suicides by gender each year in the EU

```{r}
data %>%  
  group_by(year, sex) %>% 
  summarise(suicides_no = sum(suicides_no, na.rm = T)/ sum(population, na.rm = T) * 1e5) %>% 
  ggplot(aes(year, suicides_no, color = sex)) + 
  geom_line(aes(linetype = "Worldwide")) +
  geom_line(data = data %>% 
              filter(country  %in% eu) %>% 
              group_by(year, sex) %>% 
              summarise(suicides_no = sum(suicides_no, na.rm = T)/ sum(population, na.rm = T) * 1e5), aes(linetype = "EU")) +
  scale_linetype_manual(name = "Region" ,values = c("Worldwide" = 2, "EU" = 1)) +
  labs(title = "Rate of suicide by gender", subtitle = "Since 1985 in EU and worldwide",
       x = "Year", y = "Rate of suicide")

```

### Rate of suicides by gender each year in every country

```{r}
data %>% 
  filter(country %in% eu) %>% 
  group_by(country, year, sex) %>% 
  summarise(suicides_rate = sum(suicides_no)/sum(population) * 1e5) %>% 
  ggplot(aes(year, suicides_rate, color = country)) + 
  geom_line() + facet_wrap(~sex) +
  theme(legend.position = "none") +
  labs(title = "Rate of suicide by gender", subtitle = "Since 1985 in EU",
       x = "Year", y = "Rate of suicide")

```

### Rate of suicides by age group each year in every country

```{r}
data %>% 
  filter(country %in% eu) %>% 
  group_by(country, year, age) %>% 
  summarise(suicides_rate = sum(suicides_no)/sum(population)*1e5) %>% 
  ggplot(aes(year, suicides_rate, color = country)) + 
  geom_line() + facet_wrap(~ age) +
  theme(legend.position = "none") +
  labs(title = "Rate of suicide by age", subtitle = "Since 1985 in EU",
       x = "Year", y = "Rate of suicide")

```

### Rate of suicides by age group and sex

```{r fig.height=12, fig.width=8}
data %>% 
  filter(country %in% eu) %>% 
  group_by(country, year, age, sex) %>% 
  summarise(suicides_rate = sum(suicides_no)/sum(population)*1e5) %>% 
  ggplot(aes(year, suicides_rate, color = country)) + 
  geom_line() + facet_grid(age ~ sex)+
  theme(legend.position = "none") +
  labs(title = "Rate of suicide by age and sex", subtitle = "Since 1985 in EU",
       x = "Year", y = "Rate of suicide")
```

## Finland

### Number of suicides by sex

```{r}
data %>% filter(country == "Finland") %>% 
  filter(!is.na(suicides_no)) %>% 
  ggplot() + geom_point(aes(year, suicides_no, color = fct_reorder2(age, year, suicides_no))) +
  geom_line(aes(year, suicides_no, colour = age)) + facet_wrap(~sex) +
  scale_color_discrete(name = "Age group") +
  labs(title = "Number of suicides by sex", subtitle = "in Finland",
       x = "Year", y = "Number of suicides")

```

### Rate of suicides by sex

```{r}
data %>% filter(country == "Finland") %>% 
  filter(!is.na(suicides_no)) %>% 
  ggplot() + geom_point(aes(year, suicides_rate, color = fct_reorder2(age, year, suicides_rate))) +
  geom_line(aes(year, suicides_rate, colour = age)) + facet_wrap(~sex)+
  scale_color_discrete(name = "Age group") +
  labs(title = "Rate of suicides by sex", subtitle = "in Finland",
       x = "Year", y = "Rate of suicides")

```

### Rate of suicides by generation

```{r}
data %>% filter(country == "Finland") %>% 
  filter(!is.na(suicides_no)) %>% group_by(generation, sex, year) %>% 
  summarise(suicides_rate = sum(suicides_no)/sum(population)) %>% 
  ggplot() + geom_point(aes(year, suicides_rate, color = fct_reorder2(generation, year, suicides_rate))) +
  geom_line(aes(year, suicides_rate, colour = generation)) + facet_wrap(~sex)+
  scale_color_discrete(name = "Genration") +
  labs(title = "Number of suicides by generation", subtitle = "in Finland",
       x = "Year", y = "Number of suicides")

```

