---
title: "Global terrorism"
date: 2019-05-25T10-24-00
output:
  md_document: default
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.width=12, fig.height=8,
  cache = T
)

```

# Intro

# Library

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
library(maps)
library(gganimate)
library(plotly)
library(streamgraph)
```

# Load data

```{r, cache = TRUE}
data <- read_csv("data/globalterrorismdb_0718dist.csv", col_types = cols(
  ransomamtus = col_number(),
  ransomnote = col_character(),
  attacktype3 = col_integer(),
  attacktype3_txt = col_character(),
  claimmode2 = col_integer(),
  claimmode2_txt = col_character(),
  claimmode3 = col_integer(),
  claimmode3_txt = col_character(),
  weaptype4 = col_integer(),
  weaptype4_txt = col_character(),
  weapsubtype4 = col_integer(),
  weapsubtype4_txt = col_character(),
  gname3 = col_character(),
  gsubname3 = col_character(),
  gsubname2 = col_character(),
  ransompaidus = col_number(),
  compclaim = col_integer()
))

# For world map

world <- map_data("world")
worldmap <- ggplot(data = world, aes(long, lat)) + borders("world") +
  theme(panel.background = element_blank(),
                   axis.title = element_blank(),
                   axis.line.x = element_blank(),
                   axis.ticks = element_blank(),
                   axis.text = element_blank()) +
   coord_fixed(1.2)


```

Data dimensions

```{r}
dim(data)
```

First glance

```{r}
colnames(data)
```


Deal with latitude and longtitude
```{r}
data %>% select(longitude, latitude) %>%  skim()
data %>% filter(is.na(latitude)) %>% select(location, city, provstate, country_txt, specificity) %>% filter(is.na(specificity))

data %>% map_df(~ str_detect(.,"Unknown" )) %>% summarise_all(sum, na.rm = TRUE)
```

longitude value appear to be incorrect at -8.6e+07. By compare the location with actual coordinate on google, it is obvious that a decimal separator is missing. The code below will fix that by changing it to -86.185896

```{r}
data %>% filter(longitude == min(data$longitude, na.rm = T))

data[data$eventid == 198212240004, "longitude"] <- -86.185896

```

Deal with missing longitude and latitude by using the country location as the incident location. The country location is obtained by using Mapquest API.

```{r include=FALSE}
KEY <- "YVjGrSZqSDBhUnGCcocPjsTwEpoBgaM5"
```



```{r message=FALSE, warning=FALSE}

baseurl <- "http://www.mapquestapi.com/geocoding/v1/address?"


country <- data %>%  
  filter(is.na(longitude)) %>%
  distinct(country_txt)

country <- country %>% mutate(location = str_replace_all(country_txt, " ", "+"))

country$location[country$location == "North+Yemen"] <- "Yemen"
country$location[country$location == "South+Yemen"] <- "Yemen"
country$location[country$location == "West+Germany+(FRG)"] <- "Germany"
country$location[country$location == "Yugoslavia"] <- "Croatia"
country$location[country$location == "Zaire"] <- "Democratic+Republic+of+the+Congo"
country$location[country$location == "Soviet+Union"] <- "Russia"
country$location[country$location == "West+Bank+and+Gaza+Strip"] <- "Gaza+Strip"
country$location[country$location == "Rhodesia"] <- "Zimbabwe"
country$location[country$location == "Czechoslovakia"] <- "Czech+Republic"

url <- str_c(baseurl, "key=", KEY, "&location=", country$location, "&outFormat=csv")

missing <- url %>%
  map_dfr( ~ read_csv(.) %>%
             select(Lat, Lng, GeocodeQuality, GeocodeQualityCode, Country) %>%
             filter(GeocodeQuality == "COUNTRY"))

country <- bind_cols(country,missing)

temp <- data %>%
  filter(is.na(longitude)) %>%
  select(eventid, country_txt) %>%
  left_join(country) %>%  
  select(eventid, Lng, Lat)

data <- data %>% left_join(temp, by = "eventid") %>%
  mutate(longitude = ifelse(is.na(longitude), Lng, longitude)) %>%
  mutate(latitude = ifelse(is.na(latitude), Lat, latitude))
```

Missing date and month in the data were recorded as NA, change them to 1. Then create a new column called idate which combined year, month and date together

```{r}

data$iday[data$iday == 0] <- 1
data$imonth[data$imonth == 0] <- 1

data$idate <- data %>% unite(date, iyear, imonth, iday, sep = "-") %>%
  mutate(date = lubridate::ymd(date)) %>%
  pull(date)

```

Group name contain non UTF-8 characters. The following code will fix it

```{r}
data <- data %>%
  mutate_at(vars(gname, gname2, gname3), function(x){gsub('[^ -~]', '', x)})
```

Create a label for leaflet map by making new column called popmap

```{r}
data$popmap <- data %>%
  mutate(attacktype2_txt = ifelse(is.na(attacktype2_txt)," ", str_c(", ", attacktype2_txt)),
         attacktype3_txt = ifelse(is.na(attacktype3_txt)," ", str_c(", ", attacktype3_txt)),
         weaptype2_txt = ifelse(is.na(weaptype2_txt)," ", str_c(", ", weaptype2_txt)),
         weaptype3_txt = ifelse(is.na(weaptype3_txt)," ", str_c(", ", weaptype3_txt)),
         targtype2_txt = ifelse(is.na(targtype2_txt)," ", str_c(", ", targtype2_txt)),
         targtype3_txt = ifelse(is.na(targtype3_txt)," ", str_c(", ", targtype3_txt)),
         gname2 = ifelse(is.na(gname2)," ", str_c(", ", gname2)),
         gname3 = ifelse(is.na(gname3)," ", str_c(", ", gname3))) %>%
  mutate(popmap = str_c("Country: ", country_txt, " <br/> ",
                        "Date: ", idate, " <br/> ",
                        "Attack type: ", attacktype1_txt, attacktype2_txt, attacktype3_txt, " <br/> ",
                        "Weapon: ", weaptype1_txt, weaptype2_txt, weaptype3_txt, " <br/> ",
                        "Target: ", targtype1_txt, targtype2_txt, targtype3_txt, " <br/> ",
                        "Group responsible: ", gname, gname2, gname3, " <br/> ",
                        "Casualty: ", str_replace_na(nkill), " <br/> ",
                        "Injured: ", str_replace_na(nwound), " <br/> ",
                        "Property damage: ", str_replace_na(propextent_txt))) %>% pull(popmap)
```


# Data exploration

```{r}
data <- read_csv("data/newdat.csv", col_types = cols(
  iyear = col_integer(),
  latitude = col_double(),
  longitude = col_double(),
  attacktype3_txt = col_character(),
  gname3 = col_character(),
  nkill = col_integer(),
  nwound = col_integer()
))
```

 Leaflet

```{r}
library(leaflet)

data %>%
  filter(idate == "2000-01-01") %>% 
  leaflet() %>%
      addTiles() %>%
      addMarkers(~longitude, ~latitude, label = ~data$popmap %>% purrr::map(shiny::HTML) ,
                 clusterOptions = markerClusterOptions(disableClusteringAtZoom = 12)) %>%
      setMaxBounds(~min(data$longitude), ~min(data$latitude), ~max(data$longitude), ~max(data$latitude)) %>%
      setView(0,0, 2)
```
